---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = FALSE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)

library(feck)

if (FALSE) {
# Check for functions which do not have an @examples block for roxygen
system("grep -c examples man/*Rd", intern = TRUE) |> 
  grep(":0$", x = _, value = TRUE)
}

if (FALSE) {
  covr::report(covr::package_coverage(
    line_exclusions = list()
  ))
}

if (FALSE) {
  pkgdown::build_site(override = list(destination = "../coolbutuseless.github.io/package/feck"))
}

# Makevars options to do some deep testing for CRAN

# Type conversions are sane
# PKG_FLAG=-Wconversion

# Pointer overflow checks i.e. dodgy pointer arithmetic
# PKG_CFLAGS+=-fsanitize=pointer-overflow -fsanitize-trap=pointer-overflow
# Then run in the debugger:
# R -d lldb 
# run
# testthat::test_local()

```

# feck

<!-- badges: start -->
![](https://img.shields.io/badge/cool-useless-green.svg)
[![CRAN](https://www.r-pkg.org/badges/version/feck)](https://CRAN.R-project.org/package=feck)
<!-- badges: end -->

`{feck}` is an R package for generating forward error correction codes - think 
[RAID5](https://en.wikipedia.org/wiki/Standard_RAID_levels) or [par2](https://en.wikipedia.org/wiki/Parchive) but for raw vectors.

By pre-calculating a set of *rescue blocks*, if the data is later corrupted
then these blocks can be used to repair the file.

The number of repair blocks generated is configurable depending upon user needs.

## Installation

<!-- This package can be installed from CRAN -->

<!-- ``` r -->
<!-- install.packages('feck') -->
<!-- ``` -->

You can install the latest development version from
[GitHub](https://github.com/coolbutuseless/feck) with:

``` r
# install.package('remotes')
remotes::install_github('coolbutuseless/feck')
```

<!-- Pre-built source/binary versions can also be installed from -->
<!-- [R-universe](https://r-universe.dev) -->

<!-- ``` r -->
<!-- install.packages('feck', repos = c('https://coolbutuseless.r-universe.dev', 'https://cloud.r-project.org')) -->
<!-- ``` -->

## Example: Forward Error Correction of a raw vector


```{r example}
library(feck)

# Generate some data
set.seed(1)
dat0 <- dat <- as.raw(sample(0:255, 1e4, replace = TRUE))
length(dat)

# Partitition the data into 10 chunks, and create 2 repair blocks.
# If errors occur in up to 2 of the chunks of data, then the data can 
# be repaired using these repair blocks
rblocks <- fec_prepare_raw(dat, k = 10, n = 2)
length(rblocks)

# simulate damage to data
dat[1:100] <- as.raw(0)
identical(dat, dat0)

# repair the data using the repair blocks
dat_repaired <- fec_repair_raw(dat, rblocks)
identical(dat_repaired, dat0)
```


## Example: Forward Error Correction of a file


```{r example2, error = TRUE}
library(feck)

# Same some data to a file
tmpfile <- tempfile()
saveRDS(mtcars, tmpfile)

# When reloaded it should be identical to what was saved
restore1 <- readRDS(tmpfile)
identical(mtcars, restore1)

# Create repair blocks. Consider the file split into 20 chunks,
# and create 5 repair blocks to repair errors in up to 5 chunks in the file
# The file is saved as the same as the original filename with a ".feck" suffix
fec_prepare_file(tmpfile, k = 20, n = 5)

# Corrupt the RDS file. Read it in, overwrite some bytes and save it back out
x <- readBin(tmpfile, raw(), n = file.size(tmpfile))
x[1:20] <- as.raw(0)
writeBin(x, tmpfile)

# Data is now unreadable!!
readRDS(tmpfile)

# Repair the file and save it with a ".repaired" suffix
fec_repair_file(tmpfile)

# Repaired data is now the same as the original
restore2 <- readRDS(paste0(tmpfile, ".repaired"))
identical(mtcars, restore2)
```




